---
title: "Data sets and initial insights"
author: "Marit Metz"
date: "2025-04-07"
output: html_document
---


Install packages
```{r, echo=FALSE}
library('ggplot2')
library('dplyr')
library('lme4')
library("car")
library("stringr")

```

# 1. Loading Data

Load the datasets.
##Creating the dataset from the CSV
```{r}
# Define the path to the directory containing the CSV files
# data_directory <- "..."
data_directory <- "..."
# Get a list of all .csv files in the directory
csv_files <- list.files(path = data_directory, pattern = "\\.csv$", full.names = TRUE)

# Loop through each CSV file and load it
for (file in csv_files) {
  # Read the CSV file into a data frame
  data <- read.csv(file)
  
  # Extract the file name without the path and extension
  file_name <- tools::file_path_sans_ext(basename(file)) 
  
  # Use the file name to create a variable and assign the data to it
  assign(file_name, data)
  # Optionally, store the data in a list or assign it to a variable
  # For example, storing all data frames in a list:
  data_list <- list()  # Initialize an empty list
  data_list[[file]] <- data  # Store each file's data with the file name as the list element

  print(paste("Loaded:", file))
  }
 

```



# 2. Simon task
## Selecting variables for Simon
```{r}
# Subset the dataframe
full_data_simon = simon[, c('worker_id',"exp_stage", 'rt', 'condition',  'stim_color', 'stim_side',  'correct_response','key_press','correct')]

# Update 'response' based on 'key_press'
full_data_simon$key_press <- ifelse(full_data_simon$key_press == 39, "right", 
                          ifelse(full_data_simon$key_press == 37, "left", full_data_simon$key_press))

# Update 'correct_response' based on its value
full_data_simon$correct_response <- ifelse(full_data_simon$correct_response == 39, "right", 
                                  ifelse(full_data_simon$correct_response == 37, "left", full_data_simon$correct_response))

# only keeping the test block
full_data_simon = full_data_simon[full_data_simon$exp_stage == 'test',]

#clean dataset
data_simon <- full_data_simon[, c("worker_id", "rt", "stim_color", "stim_side","condition", "key_press", "correct")]
# Rename columns
colnames(data_simon) <- c("subject", "rt", "stim_relevant", "stim_irrelevant","congruence", "R", "correct")
# Convert rt from ms to seconds
data_simon$rt <- data_simon$rt / 1000

table(data_simon$subject)
length(unique(data_simon$subject))


```

## Outlier detection used by Eisenberg et al. (2019)
Eisenberg's criteria: In general we required that median response times were longer than 200 ms, no more than 25%​ ​of responses were omitted, accuracy was higher than 60% and no single response was given more than 95% of the time.
```{r}
# Create a new column to track exclusion reasons
data_simon$exclusion_reason <- NA  # Initialize with NA for no exclusion

# Step 1: Calculate median RT per worker_id and exclude participants with median RT <= 200 ms
median_rt_per_worker <- aggregate(rt ~ subject, data = data_simon, FUN = median, na.rm = TRUE)
excluded_median_rt <- median_rt_per_worker$subject[median_rt_per_worker$rt <= 0.2]
data_simon$exclusion_reason[data_simon$subject %in% excluded_median_rt] <- "Median RT <= 200 ms"

# Step 2: Calculate the percentage of omitted responses (rt == -1) and exclude if more than 25% are omitted
omit_percentage_per_worker <- aggregate(rt ~ subject, data = data_simon, FUN = function(x) mean(x == -0.001, na.rm = TRUE))  # Calculate the proportion of omitted responses
excluded_omit <- omit_percentage_per_worker$subject[omit_percentage_per_worker$rt > 0.25]
data_simon$exclusion_reason[data_simon$subject %in% excluded_omit] <- "Omitted responses > 25%"

# Step 3: Calculate accuracy per worker_id and exclude if accuracy is <= 60%
accuracy_per_worker <- aggregate(correct ~ subject, data = data_simon, FUN = function(x) mean(x, na.rm = TRUE))  # Calculate accuracy
excluded_accuracy <- accuracy_per_worker$subject[accuracy_per_worker$correct <= 0.60]
data_simon$exclusion_reason[data_simon$subject %in% excluded_accuracy] <- "Accuracy <= 60%"

# Step 4: Check if no single response (key_press) is given more than 95% of the time for each worker
# Count the occurrences of each key_press per worker_id
response_frequency_per_worker <- table(data_simon$subject, data_simon$R)
# Convert to a data frame for easier handling
response_frequency_per_worker_df <- as.data.frame(response_frequency_per_worker)
# Rename the columns for clarity
colnames(response_frequency_per_worker_df) <- c("subject", "R", "count")

# Calculate the total number of responses per worker_id
total_responses_per_worker <- aggregate(count ~ subject, data = response_frequency_per_worker_df, FUN = sum)
# Merge the total responses with the frequency data
response_percentage_per_worker <- merge(response_frequency_per_worker_df, total_responses_per_worker, by = "subject")
# Calculate the percentage of each key_press response per worker
response_percentage_per_worker$percentage <- response_percentage_per_worker$count.x / response_percentage_per_worker$count.y

# Find workers where any key_press response is given more than 95% of the time
excluded_single_response <- response_percentage_per_worker$subject[response_percentage_per_worker$percentage > 0.95]
data_simon$exclusion_reason[data_simon$subject %in% excluded_single_response] <- "Single response > 95%"

# To create a summary of exclusions, showing the worker_id and corresponding exclusion reasons
# We will aggregate exclusion reasons for each worker_id
excluded_summary <- aggregate(exclusion_reason ~ subject, data = data_simon, FUN = function(x) paste(unique(x), collapse = ", "))

# Print the summary of excluded workers and their exclusion reasons
print(excluded_summary)
# Exclude workers in the excluded summary from the data
excluded_worker_ids <- excluded_summary$subject  # Get the list of excluded worker_ids
cat("Number of rows before exclusion:", nrow(data_simon), "\n")
# Exclude those workers from the dataset
data_simon <- data_simon[!data_simon$subject %in% excluded_worker_ids, ] # removing 218, 375, 550
# Print the number of rows before and after exclusion
cat("Number of rows before exclusion:", nrow(data_simon), "\n")

data_simon <- data_simon[, !(colnames(data_simon) %in% "exclusion_reason")]

```

## After Eisenberg's criteria
Still some very fast RTs remaining
```{r}
sum(data_simon$rt == -0.001, na.rm = TRUE)

#removing all omitted trials
data_simon = data_simon[data_simon$rt!= -0.001,]

# plotting all rt histograms based on amount of outliers
# Step 1: Calculate Z-scores for RTs
z_scores <- scale(data_simon$rt )
# Define outliers based on Z-score > 3 or < -3, or RT < 0.15
outliers_z <- abs(z_scores) > 3  # Z-score outliers
outliers_rt <- data_simon$rt < 0.15  # Fast RT outliers
outliers_combined <- outliers_z | outliers_rt
# Add the outlier information to the dataset
data_simon$outlier_combined <- outliers_combined
data_simon$outlier_z <- outliers_z
data_simon$outlier_rt <- outliers_rt
# Step 2: Count the number of outliers per participant (worker_id)
outlier_count_per_worker <- aggregate(outlier_combined ~ subject, data = data_simon, FUN = sum)
# Step 3: Sort participants by the number of outliers (in descending order)
outlier_count_per_worker <- outlier_count_per_worker[order(outlier_count_per_worker$V1, decreasing = T),]
# Step 4: Split participants into blocks of 12
outlier_count_per_worker$block <- rep(1:ceiling(nrow(outlier_count_per_worker)/12), each = 12, length.out = nrow(outlier_count_per_worker))
# Merge back to get the block information into the original data
data_simon <- merge(data_simon, outlier_count_per_worker[, c("subject", "block")], by = "subject", all.x = TRUE)
# Step 5: Plot histograms of RTs, split by block. To get an overview of the response times of each participant.
# for (i in 1:44) { #check the highest block number
#   plot_data <- data_simon[data_simon$block == i, ]
#   
#   # Create the histogram for the block
#   plot <- ggplot(plot_data, aes(x = rt)) + 
#     geom_histogram(binwidth = 0.05, fill = "black", alpha = 0.7) + 
#     xlim(0, NA) + # Customize histogram appearance
#     facet_wrap(~ subject, ncol = 3) +  # 3 histograms per row
#     labs(title = paste("RT Distribution for Block", i), 
#          x = "Response Time (RT)", 
#          y = "Frequency") + 
#     theme_minimal() + 
#     theme(strip.text = element_text(size = 8)) 
#   
#   print(plot)
# }


# Calculate the slow and fast thresholds per participant and modify RT for outliers
data_simon <- data_simon %>%
  group_by(subject) %>%
  mutate(
    mean_rt = mean(rt, na.rm = TRUE),  # Calculate the mean RT per participant
    sd_rt = sd(rt, na.rm = TRUE),      # Calculate the SD RT per participant
    slow_threshold = mean_rt + 3 * sd_rt,  # Set slow threshold (mean + 3 * SD)
    fast_threshold = 0.15,             # Set fast threshold at 0.15
    rt = ifelse(rt > slow_threshold | rt < fast_threshold, -0.001, rt)  # Set RT to -0.001 for slow or fast trials
  ) %>%
  select(-mean_rt, -sd_rt, -slow_threshold, -fast_threshold)  # Optionally remove helper columns
#removing the outliers identified above
data_simon = data_simon[data_simon$rt!= -0.001,]

# check again to see if number of omissions are below 25%
# Count the total number of rows (trials) per participant
total_trials_per_worker <- aggregate(rt ~ subject, data = data_simon, FUN = length)
# Calculate the number of omitted trials (assuming each participant should have 100 trials)
total_trials_per_worker$omitted_trials <- 100 - total_trials_per_worker$rt

# Calculate the percentage of omitted trials
total_trials_per_worker$omit_percentage <- (total_trials_per_worker$omitted_trials / 100) * 100
excluded_omit <- total_trials_per_worker$subject[total_trials_per_worker$omit_percentage > 25] # excludes participant 541
data_simon <- data_simon[!data_simon$subject %in% excluded_omit, ]
print(paste("Participants missing too much data:", paste(excluded_omit, collapse = ", ")))

#recheck median rt
median_rt_check <- aggregate(rt ~ subject, data = data_simon, FUN = median)
failed_median_rt <- median_rt_check$subject[median_rt_check$rt <= 0.2]
print(paste("Participants failing median RT check:", paste(failed_median_rt, collapse = ", ")))
#recheck accuracy
accuracy_check <- aggregate(correct ~ subject, data = data_simon, FUN = function(x) mean(x, na.rm = TRUE))
failed_accuracy <- accuracy_check$subject[accuracy_check$correct <= 0.60]
print(paste("Participants failing accuracy check:", paste(failed_accuracy, collapse = ", ")))
#recheck single response
response_freq_table <- table(data_simon$subject, data_simon$R)
response_freq_df <- as.data.frame(response_freq_table)
colnames(response_freq_df) <- c("subject", "R", "count")
# Total responses per subject
total_responses <- aggregate(count ~ subject, data = response_freq_df, sum)
response_freq_df <- merge(response_freq_df, total_responses, by = "subject")
response_freq_df$prop <- response_freq_df$count.x / response_freq_df$count.y
failed_single_response <- unique(response_freq_df$subject[response_freq_df$prop > 0.95])
print(paste("Participants overusing a single response:", paste(failed_single_response, collapse = ", ")))

# remove recheck fails
all_failed <- unique(c(failed_median_rt, failed_accuracy, failed_single_response))
data_simon <- data_simon[!data_simon$subject %in% all_failed, ]


#clean dataset
data_simon <- data_simon[, c("subject", "rt", "stim_relevant", "stim_irrelevant","congruence", "R", "correct")]
length(unique(data_simon$subject)) #518 participants remaining
# retest data has 150 of 151 participants remaining

```

##Initial effects
```{r}
### Look at factor effects on summary statistics ----
data_simon$correct <- as.numeric(as.character(data_simon$correct))
# # Create a new column 'correctness' based on the conditions 
# data_simon <- data_simon %>%
#   mutate(correctness = case_when(
#     # Conditions for congruent and stim_irrelevant == right
#     congruence == "congruent" & stim_irrelevant == "right" & 
#       ((stim_relevant == "blue" & R == "right") | (stim_relevant == "red" & R == "right")) ~ 1,
#     
#     # Conditions for congruent and stim_irrelevant == left
#     congruence == "congruent" & stim_irrelevant == "left" & 
#       ((stim_relevant == "blue" & R == "left") | (stim_relevant == "red" & R == "left")) ~ 1,
#     
#     # Conditions for incongruent and stim_irrelevant == right
#     congruence == "incongruent" & stim_irrelevant == "right" & 
#       ((stim_relevant == "blue" & R == "right") | (stim_relevant == "red" & R == "right")) ~ 0,
#     
#     # Conditions for incongruent and stim_irrelevant == left
#     congruence == "incongruent" & stim_irrelevant == "left" & 
#       ((stim_relevant == "blue" & R == "left") | (stim_relevant == "red" & R == "left")) ~ 0,
#     
#     # For other incongruent situations, set to 1
#     congruence == "incongruent" ~ 1,
#     
#     # Default case to handle situations where no other conditions match
#     TRUE ~ 0
#   ))



tapply(data_simon$rt, data_simon[,c("stim_relevant", "stim_irrrelevant", "congruence"), mean])

# effects on rt
simon_rt_factors <- lmer(rt ~ stim_relevant * congruence + (1 | subject), data = data_simon)
Anova(simon_rt_factors)
# effects on accuracy
simon_acc_factors <- glmer(correct ~ stim_relevant  * congruence + (1 | subject), data = data_simon, family = binomial("probit"))
Anova(simon_acc_factors)



# based on above significance
tapply(data_simon$rt,data_simon$congruence,mean)
tapply(data_simon$rt,data_simon$stim_relevant,mean)
tapply(data_simon$correct,data_simon$congruence,mean)
mean_rt_cong <- aggregate(rt ~ congruence + stim_relevant, data = data_simon, FUN = function(x) round(mean(x), 3))
print(mean_rt_cong)
mean_acc_cong <- aggregate(correct ~ congruence + stim_relevant, data = data_simon, FUN = function(x) round(mean(x), 3))
print(mean_acc_cong)

```

## Change data to EMC format
Above dataset doesn't allow for the function of accuracy as EMC needs it. Thus change the values of R to colors instead of sides so we can equal them to the stim_relevant values to give correctness of response. 
```{r}
simon$R <- ifelse(
  simon$correct == 1,
  simon$stim_relevant,
  ifelse(
    simon$stim_relevant == "blue", "red", "blue"
  )
)

```

To accurately define congruency in EMC the irrelevant stimulus column also needs to be defined as colors. 
```{r}
# Create or overwrite the irrelevant_stimulus column
simon$stim_irrelevant <- ifelse(
  simon$congruence == "congruent",
  simon$stim_relevant,
  ifelse(
    simon$stim_relevant == "blue", "red", "blue"
  )
)

```


Lastly selecting the required variables
```{r}


simon <- simon[, c("subject", "rt", "stim_relevant", "stim_irrelevant", "R")]
colnames(simon) <- c("subjects", "rt", "RS", "IS", "R")
simon$subjects <- as.factor(simon$subjects)
simon$R <- as.factor(simon$R)
simon$RS <- as.factor(simon$RS)
simon$IS <- as.factor(simon$IS)



```



# 3. Stroop
##Selecting variables Stroop
```{r}
# Subset the dataframe
full_data_stroop = stroop[, c('worker_id',"exp_stage", 'rt', 'condition',  'stim_color', 'stim_word',  'correct_response','key_press','correct')]

# Update 'response' based on 'key_press'
full_data_stroop$key_press <- ifelse(full_data_stroop$key_press == 66, "blue", 
                          ifelse(full_data_stroop$key_press == 82, "red", 
                                 ifelse(full_data_stroop$key_press == 71, "green", full_data_stroop$key_press)))

# Update 'correct_response' based on its value
full_data_stroop$correct_response <- ifelse(full_data_stroop$correct_response == 66, "blue", 
                                ifelse(full_data_stroop$correct_response == 82, "red", 
                                 ifelse(full_data_stroop$correct_response == 71, "green", full_data_stroop$correct_response)))
# only keeping the test block
full_data_stroop = full_data_stroop[full_data_stroop$exp_stage == 'test',]



#clean dataset
data_stroop <- full_data_stroop[, c("worker_id", "rt", "stim_color", "stim_word","condition", "key_press")]
# Rename columns
colnames(data_stroop) <- c("subject", "rt", "stim_relevant", "stim_irrelevant","congruence", "R")
# Convert rt from ms to seconds
data_stroop$rt <- data_stroop$rt / 1000
# table(data_stroop$subject)
length(unique(data_stroop$subject))

```
## Outlier detection with Eisenberg's criteria
Eisenberg's criteria: In general we required that median response times were longer than 200 ms, no more than 25%​ ​of responses were omitted, accuracy was higher than 60% and no single response was given more than 95% of the time.
```{r}
# Create a new column to track exclusion reasons
data_stroop$exclusion_reason <- NA  # Initialize with NA for no exclusion

# Step 1: Calculate median RT per worker_id and exclude participants with median RT <= 200 ms
median_rt_per_worker <- aggregate(rt ~ subject, data = data_stroop, FUN = median, na.rm = TRUE)
excluded_median_rt <- median_rt_per_worker$subject[median_rt_per_worker$rt <= 0.2]
data_stroop$exclusion_reason[data_stroop$subject %in% excluded_median_rt] <- "Median RT <= 200 ms"

# Step 2: Calculate the percentage of omitted responses (rt == -1) and exclude if more than 25% are omitted
omit_percentage_per_worker <- aggregate(rt ~ subject, data = data_stroop, FUN = function(x) mean(x == -0.001, na.rm = TRUE))  # Calculate the proportion of omitted responses
excluded_omit <- omit_percentage_per_worker$subject[omit_percentage_per_worker$rt > 0.25]
data_stroop$exclusion_reason[data_stroop$subject %in% excluded_omit] <- "Omitted responses > 25%"

# Step 3: Calculate accuracy per worker_id and exclude if accuracy is <= 60%
accuracy_per_worker <- aggregate(data_stroop$stim_relevant == data_stroop$R ~ subject, data = data_stroop, FUN = function(x) mean(x, na.rm = TRUE))  # Calculate accuracy
excluded_accuracy <- accuracy_per_worker$subject[accuracy_per_worker$correct <= 0.60]
data_stroop$exclusion_reason[data_stroop$subject %in% excluded_accuracy] <- "Accuracy <= 60%"

# Step 4: Check if no single response (key_press) is given more than 95% of the time for each worker
# Count the occurrences of each key_press per worker_id
response_frequency_per_worker <- table(data_stroop$subject, data_stroop$R)
# Convert to a data frame for easier handling
response_frequency_per_worker_df <- as.data.frame(response_frequency_per_worker)
# Rename the columns for clarity
colnames(response_frequency_per_worker_df) <- c("subject", "R", "count")

# Calculate the total number of responses per worker_id
total_responses_per_worker <- aggregate(count ~ subject, data = response_frequency_per_worker_df, FUN = sum)
# Merge the total responses with the frequency data
response_percentage_per_worker <- merge(response_frequency_per_worker_df, total_responses_per_worker, by = "subject")
# Calculate the percentage of each key_press response per worker
response_percentage_per_worker$percentage <- response_percentage_per_worker$count.x / response_percentage_per_worker$count.y

# Find workers where any key_press response is given more than 95% of the time
excluded_single_response <- response_percentage_per_worker$subject[response_percentage_per_worker$percentage > 0.95]
data_stroop$exclusion_reason[data_stroop$subject %in% excluded_single_response] <- "Single response > 95%"

# To create a summary of exclusions, showing the worker_id and corresponding exclusion reasons
# We will aggregate exclusion reasons for each worker_id
excluded_summary <- aggregate(exclusion_reason ~ subject, data = data_stroop, FUN = function(x) paste(unique(x), collapse = ", "))

# Print the summary of excluded workers and their exclusion reasons
print(excluded_summary)
# Exclude workers in the excluded summary from the data
excluded_worker_ids <- excluded_summary$subject  # Get the list of excluded worker_ids
cat("Number of rows before exclusion:", nrow(data_stroop), "\n")
# Exclude those workers from the dataset
data_stroop <- data_stroop[!data_stroop$subject %in% excluded_worker_ids, ] # removing participant 295
# Print the number of rows before and after exclusion
cat("Number of rows before exclusion:", nrow(data_stroop), "\n")

data_stroop <- data_stroop[, !(colnames(data_stroop) %in% "exclusion_reason")]

```

## Outlier investigation after removing with Eisenberg's criteria
```{r}
sum(data_stroop$rt == -0.001, na.rm = TRUE)

#removing all omitted trials
data_stroop = data_stroop[data_stroop$rt!= -0.001,]

# plotting all rt histograms based on amout of outliers
# Step 1: Calculate Z-scores for RTs
z_scores <- scale(data_stroop$rt )
# Define outliers based on Z-score > 3 or < -3, or RT < 0.15
outliers_z <- abs(z_scores) > 3  # Z-score outliers
outliers_rt <- data_stroop$rt < 0.15  # Fast RT outliers
outliers_combined <- outliers_z | outliers_rt
# Add the outlier information to the dataset
data_stroop$outlier_combined <- outliers_combined
data_stroop$outlier_z <- outliers_z
data_stroop$outlier_rt <- outliers_rt
# Step 2: Count the number of outliers per participant (worker_id)
outlier_count_per_worker <- aggregate(outlier_combined ~ subject, data = data_stroop, FUN = sum)
# Step 3: Sort participants by the number of outliers (in descending order)
outlier_count_per_worker <- outlier_count_per_worker[order(outlier_count_per_worker$V1, decreasing = T),]
# Step 4: Split participants into blocks of 12
outlier_count_per_worker$block <- rep(1:ceiling(nrow(outlier_count_per_worker)/12), each = 12, length.out = nrow(outlier_count_per_worker))
# Merge back to get the block information into the original data
data_stroop <- merge(data_stroop, outlier_count_per_worker[, c("subject", "block")], by = "subject", all.x = TRUE)
# Step 5: Plot histograms of RTs, split by block
# for (i in 1:44) { #check the highest block number
#   plot_data <- data_stroop[data_stroop$block == i, ]
#   
#   # Create the histogram for the block
#   plot <- ggplot(plot_data, aes(x = rt)) + 
#     geom_histogram(binwidth = 0.05, fill = "black", alpha = 0.7) + 
#     xlim(0, NA) + # Customize histogram appearance
#     facet_wrap(~ subject, ncol = 3) +  # 3 histograms per row
#     labs(title = paste("RT Distribution for Block", i), 
#          x = "Response Time (RT)", 
#          y = "Frequency") + 
#     theme_minimal() + 
#     theme(strip.text = element_text(size = 8)) 
#   
#   print(plot)
# }


# Calculate the slow and fast thresholds per participant and modify RT for outliers
data_stroop <- data_stroop %>%
  group_by(subject) %>%
  mutate(
    mean_rt = mean(rt, na.rm = TRUE),  # Calculate the mean RT per participant
    sd_rt = sd(rt, na.rm = TRUE),      # Calculate the SD RT per participant
    slow_threshold = mean_rt + 3 * sd_rt,  # Set slow threshold (mean + 3 * SD)
    # fast_threshold = 0.15,             # Set fast threshold at 0.15
    rt = ifelse(rt > slow_threshold , -0.001, rt)  # Set RT to -0.001 for slow or fast trials
  ) %>%
  select(-mean_rt, -sd_rt, -slow_threshold)  # Optionally remove helper columns
#removing the outliers identified above
data_stroop = data_stroop[data_stroop$rt!= -0.001,]

# check again to see if number of omissions are below 25%
# Count the total number of rows (trials) per participant
total_trials_per_worker <- aggregate(rt ~ subject, data = data_stroop, FUN = length)
# Calculate the number of omitted trials (assuming each participant should have 96 trials)
total_trials_per_worker$omitted_trials <- 96 - total_trials_per_worker$rt
# Calculate the percentage of omitted trials
total_trials_per_worker$omit_percentage <- (total_trials_per_worker$omitted_trials / 96) * 100

excluded_omit <- total_trials_per_worker$subject[total_trials_per_worker$omit_percentage > 25] 
data_stroop <- data_stroop[!data_stroop$subject %in% excluded_omit, ]

# recheck median
median_rt_check <- aggregate(rt ~ subject, data = data_stroop, FUN = median)
failed_median_rt <- median_rt_check$subject[median_rt_check$rt <= 0.2]
cat("Participants failing median RT check:", paste(failed_median_rt, collapse = ", "), "\n")
#recheck accuracy
accuracy_check <- aggregate(data_stroop$stim_relevant == data_stroop$R ~ subject, data = data_stroop, FUN = mean)
colnames(accuracy_check)[2] <- "accuracy"
failed_accuracy <- accuracy_check$subject[accuracy_check$accuracy <= 0.60]
cat("Participants failing accuracy check:", paste(failed_accuracy, collapse = ", "), "\n")
# recheck same response
response_freq_table <- table(data_stroop$subject, data_stroop$R)
response_freq_df <- as.data.frame(response_freq_table)
colnames(response_freq_df) <- c("subject", "R", "count")
# Total responses per subject
total_responses <- aggregate(count ~ subject, data = response_freq_df, sum)
response_freq_df <- merge(response_freq_df, total_responses, by = "subject")
response_freq_df$prop <- response_freq_df$count.x / response_freq_df$count.y
failed_single_response <- unique(response_freq_df$subject[response_freq_df$prop > 0.95])
cat("Participants overusing a single response:", paste(failed_single_response, collapse = ", "), "\n")


#remove
# all_fails <- unique(c(failed_median_rt, failed_accuracy, failed_single_response, perfect_accuracy))
all_fails <- unique(c(failed_median_rt, failed_accuracy, failed_single_response))
cat("Participants to exclude (any criterion):", paste(all_fails, collapse = ", "), "\n")
# Exclude from dataset
data_stroop <- data_stroop[!data_stroop$subject %in% all_fails, ]


#clean dataset
data_stroop <- data_stroop[, c("subject", "rt", "stim_relevant", "stim_irrelevant","congruence", "R")]
length(unique(data_stroop$subject)) #442 participants remaining. 

```

## Initial effects
```{r}
### Look at factor effects on summary statistics ----
# effects on rt
stroop_rt_factors <- lmer(rt ~ stim_relevant * congruence + (1 | subject), data = data_stroop)
Anova(stroop_rt_factors)
# effects on accuracy
stroop_acc_factors <- glmer(data_stroop$stim_relevant == data_stroop$R ~ stim_relevant * congruence + (1 | subject), data = data_stroop, family = binomial("probit"))
Anova(stroop_acc_factors)


# based on above significance
tapply(data_stroop$rt,data_stroop$congruence,mean)
tapply(data_stroop$rt,data_stroop$stim_relevant,mean)
tapply(data_stroop$stim_relevant==data_stroop$R,data_stroop$congruence,mean)
tapply(data_stroop$stim_relevant==data_stroop$R,data_stroop$stim_relevant,mean)

mean_rt_cong <- aggregate(rt ~ congruence + stim_relevant, data = data_stroop, FUN = function(x) round(mean(x), 3))
print(mean_rt_cong)
mean_acc_cong <- aggregate( data_stroop$stim_relevant==data_stroop$R ~ congruence + stim_relevant, data = data_stroop, FUN = function(x) round(mean(x), 3))
print(mean_acc_cong)

```


## To EMC format
```{r}

colnames(stroop) <- c("subjects", "rt", "RS", "IS", "R")
stroop$subjects <- as.factor(stroop$subjects)
stroop$R <- as.factor(stroop$R)
stroop$RS <- as.factor(stroop$RS)
stroop$IS <- as.factor(stroop$IS)


```



# 4. Attentional network task
## Selecting variables for Attentional network task
```{r}

# Subset the dataframe
full_data_ant = attention_network_task[, c('worker_id',"exp_stage", 'rt',  'flanker_type',  'cue', 'flanker_location',  'flanker_middle_direction','key_press','correct_response', 'correct')]

# Update 'response' based on 'key_press'
full_data_ant$key_press <- ifelse(full_data_ant$key_press == 39, "right", 
                          ifelse(full_data_ant$key_press == 37, "left", full_data_ant$key_press))

# Update 'correct_response' based on its value
full_data_ant$correct_response <- ifelse(full_data_ant$correct_response == 39, "right", 
                                  ifelse(full_data_ant$correct_response == 37, "left", full_data_ant$correct_response))

# only keeping the test block
full_data_ant = full_data_ant[full_data_ant$exp_stage == 'test',]




#clean dataset
data_ant <- full_data_ant[, c("worker_id", "rt", "flanker_middle_direction", "flanker_location","flanker_type", "cue",  "key_press")]
# Rename columns
colnames(data_ant) <- c("subject", "rt", "stim_relevant", "stim_irrelevant","congruence","cue", "R")
# Convert rt from ms to seconds
data_ant$rt <- data_ant$rt / 1000
# table(data_ant$subject)
length(unique(data_ant$subject))
```


## Outlier detection
Eisenberg's criteria: In general we required that median response times were longer than 200 ms, no more than 25%​ ​of responses were omitted, accuracy was higher than 60% and no single response was given more than 95% of the time.
```{r}
# Create a new column to track exclusion reasons
data_ant$exclusion_reason <- NA  # Initialize with NA for no exclusion

# Step 1: Calculate median RT per worker_id and exclude participants with median RT <= 200 ms
median_rt_per_worker <- aggregate(rt ~ subject, data = data_ant, FUN = median, na.rm = TRUE)
excluded_median_rt <- median_rt_per_worker$subject[median_rt_per_worker$rt <= 0.2]
data_ant$exclusion_reason[data_ant$subject %in% excluded_median_rt] <- "Median RT <= 200 ms"

# Step 2: Calculate the percentage of omitted responses (rt == -1) and exclude if more than 25% are omitted
omit_percentage_per_worker <- aggregate(rt ~ subject, data = data_ant, FUN = function(x) mean(x == -0.001, na.rm = TRUE))  # Calculate the proportion of omitted responses
excluded_omit <- omit_percentage_per_worker$subject[omit_percentage_per_worker$rt > 0.25]
data_ant$exclusion_reason[data_ant$subject %in% excluded_omit] <- "Omitted responses > 25%"

# Step 3: Calculate accuracy per worker_id and exclude if accuracy is <= 60%
accuracy_per_worker <- aggregate(data_ant$stim_relevant == data_ant$R ~ subject, data = data_ant, FUN = function(x) mean(x, na.rm = TRUE))  # Calculate accuracy
excluded_accuracy <- accuracy_per_worker$subject[accuracy_per_worker$correct <= 0.60]
data_ant$exclusion_reason[data_ant$subject %in% excluded_accuracy] <- "Accuracy <= 60%"

# Step 4: Check if no single response (key_press) is given more than 95% of the time for each worker
# Count the occurrences of each key_press per worker_id
response_frequency_per_worker <- table(data_ant$subject, data_ant$R)
# Convert to a data frame for easier handling
response_frequency_per_worker_df <- as.data.frame(response_frequency_per_worker)
# Rename the columns for clarity
colnames(response_frequency_per_worker_df) <- c("subject", "R", "count")

# Calculate the total number of responses per worker_id
total_responses_per_worker <- aggregate(count ~ subject, data = response_frequency_per_worker_df, FUN = sum)
# Merge the total responses with the frequency data
response_percentage_per_worker <- merge(response_frequency_per_worker_df, total_responses_per_worker, by = "subject")
# Calculate the percentage of each key_press response per worker
response_percentage_per_worker$percentage <- response_percentage_per_worker$count.x / response_percentage_per_worker$count.y

# Find workers where any key_press response is given more than 95% of the time
excluded_single_response <- response_percentage_per_worker$subject[response_percentage_per_worker$percentage > 0.95]
data_ant$exclusion_reason[data_ant$subject %in% excluded_single_response] <- "Single response > 95%"

# To create a summary of exclusions, showing the worker_id and corresponding exclusion reasons
# We will aggregate exclusion reasons for each worker_id
excluded_summary <- aggregate(exclusion_reason ~ subject, data = data_ant, FUN = function(x) paste(unique(x), collapse = ", "))

# Print the summary of excluded workers and their exclusion reasons
print(excluded_summary)
# Exclude workers in the excluded summary from the data
excluded_worker_ids <- excluded_summary$subject  # Get the list of excluded worker_ids
cat("Number of rows before exclusion:", nrow(data_ant), "\n")
# Exclude those workers from the dataset
data_ant <- data_ant[!data_ant$subject %in% excluded_worker_ids, ] # removing participants 118, 357 & 462
# Print the number of rows before and after exclusion
cat("Number of rows after exclusion:", nrow(data_ant), "\n")

data_ant <- data_ant[, !(colnames(data_ant) %in% "exclusion_reason")]
length(unique(data_ant$subject))

```



## After Eisenberg's criteria
```{r}
sum(data_ant$rt == -0.001, na.rm = TRUE)

#removing all omitted trials
data_ant = data_ant[data_ant$rt!= -0.001,]

# plotting all rt histograms based on amout of outliers
# Step 1: Calculate Z-scores for RTs
z_scores <- scale(data_ant$rt )
# Define outliers based on Z-score > 3 or < -3, or RT < 0.15
outliers_z <- abs(z_scores) > 3  # Z-score outliers
outliers_rt <- data_ant$rt < 0.15  # Fast RT outliers
outliers_combined <- outliers_z | outliers_rt
# Add the outlier information to the dataset
data_ant$outlier_combined <- outliers_combined
data_ant$outlier_z <- outliers_z
data_ant$outlier_rt <- outliers_rt
# Step 2: Count the number of outliers per participant (worker_id)
outlier_count_per_worker <- aggregate(outlier_combined ~ subject, data = data_ant, FUN = sum)
# Step 3: Sort participants by the number of outliers (in descending order)
outlier_count_per_worker <- outlier_count_per_worker[order(outlier_count_per_worker$V1, decreasing = T),]
# Step 4: Split participants into blocks of 12
outlier_count_per_worker$block <- rep(1:ceiling(nrow(outlier_count_per_worker)/12), each = 12, length.out = nrow(outlier_count_per_worker))
# Merge back to get the block information into the original data
data_ant <- merge(data_ant, outlier_count_per_worker[, c("subject", "block")], by = "subject", all.x = TRUE)
# Step 5: Plot histograms of RTs, split by block
# for (i in 1:44) { #check the highest block number
#   plot_data <- data_ant[data_ant$block == i, ]
#   
#   # Create the histogram for the block
#   plot <- ggplot(plot_data, aes(x = rt)) + 
#     geom_histogram(binwidth = 0.05, fill = "black", alpha = 0.7) + 
#     xlim(0, NA) + # Customize histogram appearance
#     facet_wrap(~ subject, ncol = 3) +  # 3 histograms per row
#     labs(title = paste("RT Distribution for Block", i), 
#          x = "Response Time (RT)", 
#          y = "Frequency") + 
#     theme_minimal() + 
#     theme(strip.text = element_text(size = 8)) 
#   
#   print(plot)
# }


# Calculate the slow and fast thresholds per participant and modify RT for outliers
data_ant <- data_ant %>%
  group_by(subject) %>%
  mutate(
    mean_rt = mean(rt, na.rm = TRUE),  # Calculate the mean RT per participant
    sd_rt = sd(rt, na.rm = TRUE),      # Calculate the SD RT per participant
    slow_threshold = mean_rt + 3 * sd_rt,  # Set slow threshold (mean + 3 * SD)
    fast_threshold = 0.15,             # Set fast threshold at 0.15
    rt = ifelse(rt > slow_threshold | rt < fast_threshold, -0.001, rt)  # Set RT to -0.001 for slow or fast trials
  ) %>%
  select(-mean_rt, -sd_rt, -slow_threshold, -fast_threshold)  # Optionally remove helper columns
#removing the outliers identified above
data_ant = data_ant[data_ant$rt!= -0.001,]

# check again to see if number of omissions are below 25%
# Count the total number of rows (trials) per participant
total_trials_per_worker <- aggregate(rt ~ subject, data = data_ant, FUN = length)
# Calculate the number of omitted trials (assuming each participant should have 96 trials)
total_trials_per_worker$omitted_trials <- 144 - total_trials_per_worker$rt
# Calculate the percentage of omitted trials
total_trials_per_worker$omit_percentage <- (total_trials_per_worker$omitted_trials / 144) * 100
excluded_omit <- total_trials_per_worker$subject[total_trials_per_worker$omit_percentage > 25] # excludes participant 305
data_ant <- data_ant[!data_ant$subject %in% excluded_omit, ]


# Recheck: Median RT > 200 ms
median_rt_check <- aggregate(rt ~ subject, data = data_ant, FUN = median)
failed_median_rt <- median_rt_check$subject[median_rt_check$rt <= 0.2]
cat("Participants failing median RT check:", paste(failed_median_rt, collapse = ", "), "\n")

# Recheck: Accuracy > 60%
accuracy_check <- aggregate(data_ant$stim_relevant == data_ant$R ~ subject, data = data_ant, FUN = mean)
colnames(accuracy_check)[2] <- "accuracy"
failed_accuracy <- accuracy_check$subject[accuracy_check$accuracy <= 0.60]
cat("Participants failing accuracy check:", paste(failed_accuracy, collapse = ", "), "\n")

# Recheck: No single response > 95%
response_freq_table <- table(data_ant$subject, data_ant$R)
response_freq_df <- as.data.frame(response_freq_table)
colnames(response_freq_df) <- c("subject", "R", "count")

total_responses <- aggregate(count ~ subject, data = response_freq_df, sum)
response_freq_df <- merge(response_freq_df, total_responses, by = "subject")
response_freq_df$prop <- response_freq_df$count.x / response_freq_df$count.y

failed_single_response <- unique(response_freq_df$subject[response_freq_df$prop > 0.95])
cat("Participants overusing a single response:", paste(failed_single_response, collapse = ", "), "\n")


# Exclude all failing participants
# all_fails <- unique(c(failed_median_rt, failed_accuracy, failed_single_response, perfect_accuracy))
all_fails <- unique(c(failed_median_rt, failed_accuracy, failed_single_response))
cat("Participants to exclude (any criterion):", paste(all_fails, collapse = ", "), "\n")

# Final exclusion
data_ant <- data_ant[!data_ant$subject %in% all_fails, ]


#clean dataset
data_ant <- data_ant[, c("subject", "rt", "stim_relevant", "stim_irrelevant","congruence","cue", "R")]
length(unique(data_ant$subject)) # 434 participants remaining

```


## Initial effects
```{r}
### Look at factor effects on summary statistics ----
# effects on rt
ant_rt_factors <- lmer(rt ~ stim_relevant * stim_irrelevant* congruence * cue + (1 | subject), data = data_ant)
Anova(ant_rt_factors)
# effects on accuracy
ant_acc_factors <- glmer(data_ant$stim_relevant == data_ant$R ~ stim_relevant *stim_irrelevant * congruence * cue + (1 | subject), data = data_ant, family = binomial("probit"))
Anova(ant_acc_factors)




# based on above significance
tapply(data_ant$rt,data_ant$congruence,mean)
tapply(data_ant$rt,data_ant$stim_relevant,mean)
tapply(data_ant$rt,data_ant$cue,mean)
tapply(data_ant$rt,data_ant$stim_irrelevant,mean)


tapply(data_ant$stim_relevant==data_ant$R,data_ant$congruence,mean)
tapply(data_ant$stim_relevant==data_ant$R,data_ant$stim_irrelevant,mean)
tapply(data_ant$stim_relevant==data_ant$R,data_ant$cue,mean)

mean_rt_cong <- aggregate(rt ~ congruence + stim_relevant + stim_irrelevant + cue, data = data_ant, FUN = function(x) round(mean(x), 3))
print(mean_rt_cong)
mean_acc_cong <- aggregate( data_ant$stim_relevant==data_ant$R ~ congruence + stim_relevant + stim_irrelevant + cue, data = data_ant, FUN = function(x) round(mean(x), 3))
print(mean_acc_cong)
```


## To EMC format
```{r}

ant$stim_flanker <- ifelse(
  ant$congruence == "congruent",
  ant$stim_relevant,
  ifelse(
    ant$congruence == "incongruent",
    ifelse(ant$stim_relevant == "left", "right", "left"),
    'neutral'  # or "neutral", depending on what you want for neutral trials
  )
)
ant <- ant[, c("subject", "rt", "stim_relevant","stim_flanker", "stim_irrelevant","cue", "R")]
colnames(ant) <- c("subjects", "rt", "RS", "IS","IS_loc","CUE", "R")
ant$subjects <- as.factor(ant$subjects)
ant$R <- as.factor(ant$R)
ant$RS <- as.factor(ant$RS)
ant$IS <- as.factor(ant$IS)
ant$IS_loc <- as.factor(ant$IS_loc)
ant$CUE <- as.factor(ant$CUE)


```



# 5. Saving the datafiles
## A specific file
```{r}
# Define the path to your desired location
file_path <- ".../data_simon.Rdata"

# Save the dataset to the specified location
save(loc_glob, file = file_path)

```


## All data objects
```{r}
# List all objects in the workspace
all_objects <- ls()

# Filter objects that start with 'data_' but exclude 'data_directory' and 'data_list'
data_objects <- all_objects[grepl("^data_", all_objects) & !all_objects %in% c("data_directory", "data_list", "data_objects")]

# Define the directory where the files should be saved
save_directory <- ".../Data/"

# Save each object individually
for (obj in data_objects) {
  # Define the file path for each object
  file_path <- paste0(save_directory, obj, ".RData")
  
  # Save the object to the specified path
  save(list = obj, file = file_path)
}

```



